Fast and exact visibility on digitized shapes and application to saliency-aware normal estimation

### 2025-04-25 DGMM2025/computation.tex
Working on 3

@@

A \emph{lattice map along axis $j$} is a set of pairs $(q,L_q)$, with
$q \in \Z^{d-1}$ and $L_q \in \mathbb{L}$. The lattice map $M_K$
\emph{represents} $K$ when, for any point $p \in K$, the pair
$(\pi_j(p), L_{\pi_j(p)})$ exists in $M_K$ and $p_j \in L_{\pi_j(p)}$,
with $p_j$ the $j$-th coordinate of $p$, and reciprocally, $\forall
(q,L_q) \in M_K, \forall i \in L_q, q+i\ve_j \in K$.

To improve computation time, we will use lattice maps of integer
intervals.

\begin{definition}
  A sequence of intervals is an ordered list of intervals $L = ([a_i,b_i])_{i=1,\ldots,n}$ such that $b_i + 1 < a_{i+1}$, i.e., disjoint intervals with at least a missing integer.
\end{definition}

\begin{definition}
  A translation of an interval $A$ is defined as $A+t \coloneqq [a+t, b+t]$
\end{definition}

\begin{definition}
  A translation of a sequence of intervals $L$ is the sequence defined as $L+t \coloneqq \{L_1+t,\ldots,L_n+t\}$
\end{definition}

\begin{definition}
  All translations $T$ of an interval or a sequence of intervals $A$ in a sequence of intervals $L$ are defined as $ T \coloneqq \{ t, st. A+t \subset L\}$
\end{definition}

We will consider lattice maps as pairs of (shift, intervals), where shifts are points $p \in \R^{d-1}$. All coordinates are doubled, so that the represented cell $\forall c \in \R^d$ will have a dimension equal to $\sum_{i=1}^d \left(c[i]\mod d\right)$.
In figure~\ref{fig:lattice-representation}, we display a representation of lattice maps applied to a 2-d cell complex where all coordinates are already doubled. Note that the examples are in 2-d, the shown results are in 3-d and the algorithm is for n-d.

\begin{figure}
  \centering
  \input{lattice.tikz}
  \caption{\label{fig:lattice-representation} Representation of a
    2d cell complex (here the star of a given curve) as a lattice
    map. From 65 cells (with two coordinates) on the left, we get 11
    intervals on the right.}
\end{figure}

@@

\begin{algorithm}
  \caption{Given a subcomplex $C$ and an integer $r$, returns the visibility from every point of $C$ up to distance $r$. The main axis is supposed to be $z$, while $x,y$ are the auxiliary axes.}
  \label{alg:visibility}
  \begin{algorithmic}
    \Function{Visibility}{$C$: Subcomplex, $r$: Integer}
    \State $\Omega \gets \Call{Star}{C}$ \Comment{Lattice map of the star of the studied cell complex}
    \State $Directions \gets \Call{GetAllPrimalDirections}{r}$
    \State $V: \text{vector of boolean} \gets [0, \ldots, 0]$ \Comment{length $Size(Directions) \times \#C.pointels$}
    \State $low, high \gets \Call{BoundingBoxZ}{\Omega}$
    \ForAll{$d$ in $Directions$}
    \ForAll{shift $S$ in $\Omega$}
    \State $R \gets [low, high]$
    \ForAll{pair $P$ in $\Call{Star}{\text{d}}$}
    \State $R \gets R \cap \Call{Translations}{P.intervals,\Omega[S + P.shift]}$
    \EndFor
    \State \Call{UpdateVisibility}{$V$, $R$}
    \EndFor
    \EndFor
    \State \Return $V$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

In order to compute the visibility, we first compute $\Omega = \text{Star}(C)$. Then for each primal direction of
coordinates at most $r$, we look at all the possible positions this specific direction does link 2 visible points
of the cell complex. In order to do so, for each shift in $\Omega$, we compute the positions where points are
visible using the same method as presented in~\ref{fig:visibility-algorithm-evolution}. Some results of visible
points are present in figure~\ref{fig:visibility-results}.

@@

To efficiently compute the intersection of 2 lists of intervals $K$ and $L$, we go through the first 2 unvisited
elements of the 2 lists. If they don't overlap (i.e.\ the start and end of the 1st interval are both smaller than
the start of the other), then we can discard the smaller interval. Else, we construct the intersection of these
2 intervals as one of the resulting intervals. We then can skip the interval with the smallest end. If both
intervals have the same end, then both intervals can be skipped. Doing this computation until one of the lists
is empty returns the intersection of both lists of intervals.

@@

\begin{figure}
  \centering
  \begin{tabular}{c c}
    \includegraphics[width=0.4\textwidth]{pictures/visibility_from_given_point_r_10} &
    \includegraphics[width=0.4\textwidth]{pictures/visibility_aware_of_features}
  \end{tabular}
  \caption{Examples of visibility on a fandisk and a torus knot. First one is on the edge
    of a fandisk, where the visibility stops at the edges. The other one is in the interior
    of a torus knot, where we can see that the visibility is feature-aware and doesn't see
    the other side of the torus' branch.}
  \label{fig:visibility-results}
\end{figure}
%
%    \begin{figure}
%        \begin{center}
%            \includegraphics[width=0.8\textwidth]{pictures/visibility_from_given_point_r_10}
%            \caption{Visibility of a point on an edge of a fandisk}
%            \label{fig:visibility-fandisk}
%        \end{center}
%    \end{figure}
%    \begin{figure}
%        \begin{center}
%            \includegraphics[width=0.8\textwidth]{pictures/visibility_aware_of_features}
%            \caption{Visibility of a point on a torus knot (the visibility is feature-aware)}
%            \label{fig:visibility-torus-knot}
%        \end{center}
%    \end{figure}

@@

\begin{figure}[t]
  \input{pointelscomptime.tikz}
  \caption{Computation Time of Visibility as a Function of pointels amount. We compare the computation time of
    the naive breadth first visibility algorithm against our version using intervals. We input the same radius to
    both versions (10, 20, 30) and test it on 6 different figures (goursat, torus, rcube, sphere9, leopold, d20)
    with various gridsteps, giving us figures with pointel amounts ranging from 520 to 390235.}
  \label{fig:meanvisibility-computationComplexity}
\end{figure}

>>>>

\begin{figure}[t]
  \centering
  \input{lattice.tikz}
  \caption{\label{fig:lattice-representation} Representation of a
    2d cell complex (here the star of a given curve) as a lattice
    map. From 65 cells (with two coordinates) on the left, we get 11
    intervals on the right.}
\end{figure}

@@

A \emph{lattice map along axis $j$} is a set of pairs $(q,L_q)$, with
$q \in \Z^{d-1}$ and $L_q \in \mathbb{L}$. The lattice map $M_K$
\emph{represents} $K$ when, for any point $p \in K$, the pair
$(\pi_j(p), L_{\pi_j(p)})$ exists in $M_K$ and $p_j \in L_{\pi_j(p)}$,
with $p_j$ the $j$-th coordinate of $p$, and reciprocally, $\forall
(q,L_q) \in M_K, \forall i \in L_q, q+i\ve_j \in K$. A lattice map is
thus made of pairs $(q,L_q)$, where $q$ is the \emph{shift} while
$L_q$ is the \emph{intervals}. For a point $q$ which is a shift of
$M_K$, we write $M_K[q]$ for the corresponding intervals $L_q$.

The global algorithm for computing visibility is given in
Algorithm~\ref{alg:visibility}.  Given a digital input surface $C$ (a
subcomplex), we compute the lattice map $\Omega$ of $\Star{C}$ once,
generally along the axis where $C$ is the most elongated. In order to
compute visibility up to a given distance $r$, we consider all non
null primitive vectors $D$ with infinity norm no greater than $r$.
For each vector $\vv$, we compute the lattice map $M_\vv$ of its
intersected cells (as a vector from $\mathbf{0}$ to $\vv$ in
$\R^d$). The visibility along the direction $\vv$ reduces to computing
for each pair of $M_\vv$ the possible translations of intervals in
$\Omega$, and then intersecting all the translations.

%% \begin{definition}
%%   A sequence of intervals is an ordered list of intervals $L = ([a_i,b_i])_{i=1,\ldots,n}$ such that $b_i + 1 < a_{i+1}$, i.e., disjoint intervals with at least a missing integer.
%% \end{definition}

%% \begin{definition}
%%   A translation of an interval $A$ is defined as $A+t \coloneqq [a+t, b+t]$
%% \end{definition}

%% \begin{definition}
%%   A translation of a sequence of intervals $L$ is the sequence defined as $L+t \coloneqq \{L_1+t,\ldots,L_n+t\}$
%% \end{definition}

%% \begin{definition}
%%   All translations $T$ of an interval or a sequence of intervals $A$ in a sequence of intervals $L$ are defined as $ T \coloneqq \{ t, st. A+t \subset L\}$
%% \end{definition}

%% We will consider lattice maps as pairs of (shift, intervals), where shifts are points $p \in \R^{d-1}$. All coordinates are doubled, so that the represented cell $\forall c \in \R^d$ will have a dimension equal to $\sum_{i=1}^d \left(c[i]\mod d\right)$.
%% In figure~\ref{fig:lattice-representation}, we display a representation of lattice maps applied to a 2-d cell complex where all coordinates are already doubled. Note that the examples are in 2-d, the shown results are in 3-d and the algorithm is for n-d.

\begin{algorithm}
  \caption{Given a subcomplex $C$ and an integer $r$, returns the visibility from every point of $C$ up to distance $r$. The main axis is supposed to be $z$, while $x,y$ are the auxiliary axes.}
  \label{alg:visibility}
  \begin{algorithmic}
    \Function{Visibility}{$C$: Subcomplex, $r$: Integer}
    \State $\Omega \gets M_{\Call{Star}{C}}$ \Comment{Lattice map of the star of input subcomplex}
    \State $Directions \gets \Call{GetAllPrimalDirections}{r}$
    \State $V: \text{vector of boolean} \gets [0, \ldots, 0]$ \Comment{length $Size(Directions) \times \#C.pointels$}
    \State $low, high \gets \Call{BoundingBoxZ}{\Omega}$
    \ForAll{$\vv$ in $Directions$}
    \ForAll{shift $S$ in $\Omega$}
    \State $R \gets [low, high]$
    \ForAll{pair $P$ in $M_{\Call{Star}{[\mathbf{0},\mathbf{v}]}}$}
    \State $R \gets R \cap \Call{Translations}{P.intervals,\Omega[S + P.shift]}$
    \EndFor
    \State \Call{UpdateVisibility}{$V$, $R$}
    \EndFor
    \EndFor
    \State \Return $V$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

Figure~\ref{fig:visibility-algorithm-evolution} shows an example of
execution of this algorithm in an elementary case. It reduces to
determining all the possible inclusions by translation of sequences of
intervals in another sequence of intervals (note that for
pairwise visibility, it is enough to consider all the possible
inclusions by translation of one interval in another sequence of
intervals), and then intersecting progressively the results. Of
course, the process is stopped as soon as the intersection is
empty. Some results of visible points are displayed on
figure~\ref{fig:visibility-results} on 3d shapes.

%% In order to compute the visibility, we first compute $\Omega = \text{Star}(C)$. Then for each primal direction of
%% coordinates at most $r$, we look at all the possible positions this specific direction does link 2 visible points
%% of the cell complex. In order to do so, for each shift in $\Omega$, we compute the positions where points are
%% visible using the same method as presented in~\ref{fig:visibility-algorithm-evolution}. Some results of visible
%% points are present in figure~\ref{fig:visibility-results}.

@@

To efficiently compute the intersection of two sequences of intervals
$K$ and $L$ (Algorithm~\ref{alg:intersection}), we go through the
first $2$ unvisited elements of the $2$ lists. If they do not overlap
(i.e.\ the start and end of the 1st interval are both smaller than the
start of the other), then we can discard the smaller interval. Else,
we construct the intersection of these 2 intervals as one of the
resulting intervals. We then can skip the interval with the smallest
end. If both intervals have the same end, then both intervals can be
skipped. Doing this computation until one of the lists is empty
returns the intersection of both lists of intervals.

@@

\begin{figure}
  \centering
  \begin{tabular}{c c}
    \includegraphics[width=0.4\textwidth]{pictures/visibility_from_given_point_r_10} &
    \includegraphics[width=0.4\textwidth]{pictures/visibility_aware_of_features}
  \end{tabular}
  \caption{Examples of visibilities from a source point:
    (left) the visibility is stopped at the sharp edge, (right) the
    visibility does not cross the gap between the two parts.}
  \label{fig:visibility-results}
\end{figure}
%
%    \begin{figure}
%        \begin{center}
%            \includegraphics[width=0.8\textwidth]{pictures/visibility_from_given_point_r_10}
%            \caption{Visibility of a point on an edge of a fandisk}
%            \label{fig:visibility-fandisk}
%        \end{center}
%    \end{figure}
%    \begin{figure}
%        \begin{center}
%            \includegraphics[width=0.8\textwidth]{pictures/visibility_aware_of_features}
%            \caption{Visibility of a point on a torus knot (the visibility is feature-aware)}
%            \label{fig:visibility-torus-knot}
%        \end{center}
%    \end{figure}

@@

\begin{figure}[t]
  \input{pointelscomptime.tikz}
  \caption{Computation time of visibility as a function of the number
    of pointels. We compare the computation time of the naive breadth
    first visibility algorithm against our algorithm using
    intervals. We input the same maximal radius to both versions (10,
    20, 30) and test it on 6 different figures (goursat, torus, rcube,
    sphere9, leopold, d20) with various gridsteps, giving us figures
    with pointel amounts ranging from 520 to 390235.}
  \label{fig:meanvisibility-computationComplexity}
\end{figure}
